Type Refinements for the Working Class

### 2015-12-19 paper.tex
finish definitions of type hood, value typehood

%% \begin{thm}
%%   For any refinement $\Refines{\Upsilon}{\phi}{\tau}$, from
%%   $\IsValueType{\Upsilon}{\phi}{\tau}$ we may conclude
%%   $\IsType{\Upsilon}{\phi}{\tau}$.
%% \end{thm}
%% \begin{proof}
%%   Fix a renaming
%%   \begin{tikzcd}[cramped]
%%     \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
%%   \end{tikzcd}%
%%   , terms $\Member{L,M,N}{\Exprs_\tau(\Upsilon')}$ such that
%%   $\SBinRel{\sim_\tau}{\Upsilon'}{L}{M}$ and $\RMemEq{\Upsilon'}{M}{N}{\phi}$;
%%   by hypothesis, we also have some
%%   $\Of{\psi}{\Hom{\Sets^\SCtx}{\Yoneda{\Upsilon}}{\PERs{\Values_\tau^2}}}$,
%%   factoring through $\phi$.
%%
%%   We need to show that $\RMemEq{\Upsilon'}{L}{N}{\phi}$.
%%
%% \end{proof}

>>>>

Now, let us return briefly to Martin-L\"of's notion of type, which we call a
\emph{value type}. The minimal data of such a type's definition consists in a
partial equivalence relation on values,
$\Of{\phi}{\Hom{\Sets^\SCtx}{\Yoneda{\Upsilon}}{\PERs{\Values_\tau}}}$.
It is easy to turn such a relation into a computational type
$\IsType{\Upsilon}{\VToCType(\phi)}{\tau}$, as follows:
\[
  \begin{tikzcd}[sep=huge]
    \IMode{\Yoneda{\Upsilon}\times\Values_\tau^2}
      \arrow[r, "\IMode{\lambda(\phi)}"]
      \arrow[d, swap, "\IMode{\Pair{\ArrId{}}{\iota}}"]
      &
    \IMode{\Omega}\\
    \IMode{\Yoneda{\Upsilon}\times\Exprs_\tau^2}
    \arrow[ur, swap, dotted, "\OMode{\lambda(\VToCType(\phi))}"]
  \end{tikzcd}
\]
where $\VToCType(\phi)$ is the computational extension of $\phi$ in the following sense:
\begin{align*}
  \ADefine{
    \VToCType(\phi)(M,N)
  }{
    \forall M', N' \in \Values_\tau.\
    M\sim_\tau M' \land N\sim_\tau N' \implies \phi(M',N')
  }
\end{align*}

\begin{thm}
When viewed as an internal functor
$\PERs{\Values_\tau}\to\PERs{\Exprs_\tau}$, $\VToCType$ is the left adjoint to
the canonical inclusion
$\Of{\iota}{\PERs{\Exprs_\tau}\inj\PERs{\Values_\tau}}$.
\end{thm}
\begin{proof}
  We will exhibit a counit $\Of{\epsilon}{\VToCType\circ\iota\to
  \ArrId{\PERs{\Values_\tau}}}$ and a unit
  $\Of{\eta}{\ArrId{\PERs{\Exprs_\tau}}\to\iota\circ\VToCType}$ for
  the adjunction $\IMode{\VToCType}\dashv\IMode{\iota}$.

  The counit witnesses the fact that for any refinement
  $\Of{\phi}{\PERs{\Exprs_\tau}}$ and terms $\Member{M,N}{\Exprs_\tau}$, if
  $\VToCType(\iota(\phi))(M,N)$, then $\phi(M,N)$; this is evidently the case,
  because the premise only obtains if $M,N$ are computationally equal to
  values.  The unit witnesses the fact that for any value PER
  $\Of{\psi}{\PERs{\Values_\tau}}$ and values $\Member{M,N}{\Values_\tau}$, if
  $\psi(M,N)$, then $\iota(\VToCType(\psi))(M,N)$; this is clearly true, since
  $\VToCType(\psi)$ can do nothing but coarsen $\psi$'s relation on values.

\end{proof}

\begin{definition}[Closure]
  For any computational type $\IsType{\Upsilon}{\phi}{\tau}$, let
  $\Define{\underline{\phi}}{\VToCType(\iota(\phi))}$ be called $\phi$'s
  \emph{closure}, which we will use shortly in order to give a formal definition
  to the notion of value types.
\end{definition}

\begin{definition}[Value Types]
  We can give a simple characterization of Martin-L\"of's \emph{value types} in
  terms of our computation types. For any type $\IsType{\Upsilon}{\phi}{\tau}$,
  we will say $\IsValueType{\Upsilon}{\phi}{\tau}$ in case $\phi$ is equal to
  its closure, i.e.\ $\EqRefines{\Upsilon}{\phi}{\underline{\phi}}{\tau}$.
\end{definition}
